\bf ``A p-value of $<$ 0.05 was considered statistically significant'': a content analysis of published statistical methods sections

### 2021-06-29 manuscript/asa_template/asa_template_almostdone.tex
update manuscript files

All \emph{PLOS ONE} articles are freely accessible via the PLOS
Application Programming Interface (API). This enabled us to conduct
searches of full-text articles and analyse data on articles' text
content and general attributes such as publication date and field(s) of
research. Since \emph{PLOS ONE} does not prescribe a set article format, we applied a two-step
approach to identify statistical methods sections:

@@

\emph{Step 2}: \emph{PLOS ONE} does not use standardised headings
to preface statistical methods sections. To address this, we performed
partial matching on available headings against frequently used terms in
initial search results: `Statistical analysis', `Statistical analyses',
`Statistical method', `Statistics', `Data analysis' and `Data analyses'.

Data were downloaded on 3 July 2020. For
records that did not pass Step 2, we reviewed where initial search terms
appeared in the full-text, to estimate the proportion of
statistical methods sections that were missed under our chosen approach.

@@

Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. 
Frequent words related to the largest topic (topic 2, n = 1834) reflected descriptions of sample size calculations
 and represented 19$\%$ of studies analysed. Other topics indicated pilot studies (topic 5, n = 834),
safety/tolerability studies (topic 6, n = 524), descriptive statistics (topic 7, n = 333),
intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852). A review of word counts by topic 
identified statistical methods sections that were only one word, including ``ANOVA'',
``t-test'', ``SPSS'' and even ``SSPS''.

@@

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
total number of API search results peaked at volumes 8 (n = 19,045) and
9 (n = 19,045) in years 2013 and 2014. This trend aligned with the total
number of papers published in \emph{PLOS ONE} over the same period. The
percentage of records that included a statistical methods section varied between 64\%
(volume 2) and 86\% (volume 9). All DOIs included ``Biology and life
sciences'' (n = 107,584), ``Earth sciences'' (n = 7,605) and/or ``Computer and
information sciences'' (n = 5,190) in their top 3 subject classifications
(Figure \ref{fig:plos-n}B).

@@

For studies excluded based on section headings (n = 20,116), analysis-related search terms for 2,136 DOIs appeared in the general text of the Materials and Methods section. For studies without a matching Materials and Methods section,
 search terms oftern appeared in the Introduction (n = 728) and Results and/or Discussion (n = 981) sections. 4,010 DOIs included at least 1 section heading that contained the terms ``analysis'' or ``analyses'' without direct reference to statistical method(s) (e.g. Microarray analysis).

@@

\caption{\label{fig:plos-clouds}Word clouds for ten topics for statistical methods sections published in PLOS ONE}\label{fig:unnamed-chunk-6}
\end{figure}

>>>>

All \emph{PLOS ONE} articles are freely accessible via the PLOS
Application Programming Interface (API). This enabled us to conduct
searches of full-text articles and analyse data on articles' text
content and general attributes such as publication date and field(s) of
research. We applied a two-step
approach to identify statistical methods sections:

@@

\emph{Step 2}: \emph{PLOS ONE} does not use standardised headings
to preface statistical methods sections. To address this, we performed
partial matching on available headings against frequently used terms in
initial search results: `Statistical analysis', `Statistical analyses',
`Statistical method', `Statistics', `Data analysis' and `Data analyses'. Data were downloaded on 3 July 2020.

@@

Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. 
Words related to the largest topic (topic 2, n = 1,834) reflected descriptions of sample size calculations, for example, ``sample'',``power'' and ``80'' (i.e. $80\%$ power) . Other topics indicated pilot studies (topic 5, n = 834),
safety/tolerability studies (topic 6, n = 524), descriptive statistics (topic 7, n = 333),
intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852). A review of word counts by topic 
identified statistical methods sections that were only one word, including ``ANOVA'',
``t-test'', ``SPSS'' and even ``SSPS''.

@@

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
total number of API search results peaked at volumes 8 (n = 19,045) and
9 (n = 19,045) in years 2013 and 2014.  The
percentage of records that included a statistical methods section varied between 64\%
(volume 2) and 86\% (volume 9). All DOIs included ``Biology and life
sciences'' (n = 107,584), ``Earth sciences'' (n = 7,605) and/or ``Computer and
information sciences'' (n = 5,190) in their top 3 subject classifications
(Figure \ref{fig:plos-n}B).

@@

%For studies excluded based on section headings (n = 20,116), analysis-related search terms for 2,136 DOIs appeared in the general text of the Materials and Methods section. In other cases,
 %search terms oftern appeared in the Introduction (n = 728) and Results and/or Discussion (n = 981) sections. 4,010 DOIs included at least 1 section heading that contained the terms ``analysis'' or ``analyses'' without direct reference to statistical method(s) (e.g. Microarray analysis).

@@

\caption{\label{fig:plos-clouds}Word clouds for ten topics for statistical methods sections published in PLOS ONE}\label{fig:plos-clouds}
\end{figure}
